### 2020-08-21 vertical-thermal-review.Rmd
bit of rearranging of content on species-level traits

*Isoprene production*
-  Emission as a species trait- Tyeen Taylor, Marielle Smith 
- The high temperature carbon compensation point of sun leaves is higher in isoprene emitting species than in non-emitting species, so it is to be expected that other aspects of heat tolerance will be similarly affected by isoprenes (https://doi.org/10.1111/pce.13564)
-  Within species, isporene production scales with light/ T

*Deciduous leaf habit* 
- (Meakem et al. 2018) and refs therein

@@

**Leaf thermal tolerance (Tcrit/ T50)**
- define.
Typical values ~45-50$\circ$C, varying somewhat 
across latitude/climate (O’sullivan et al. 2017),
with elevation [@feeley_thermal_2020;  Slot unpublished data],
with leaf traits [Sastry et al. 2018, @zhang_photosynthetic_2012].

@@

**Differences across forest vertical gradients in biophysical conditions, plant traits, and metabolism and its thermal responses scale up to affect whole-tree ecology in several ways.** First, vertical gradients in the biophysical environment shape which plant strategies, or sets of traits, are competitive in understory versus canopy conditions. 
-	*Spring and fall leaf phenology in temperate deciduous forests (Augsburger). 	Mechanism: more moderate conditions (T, wind) in understory (REF).*
*Among species that can be deciduous, greater proportion of deciduous individuals in larger size classes (Condit et al. 2000). One potential mechanism is lower temperature/hydraulic stress in understory*

>>>>

*Isoprene production*
-  Isoprene production *capability* as a trait (actual isoprene production is in section 4. I'm not sure whether Ty's content belongs here or in that section)

@@

**Isoprene production**
- high emission rates occur in the mid-canopy, and even from very small statured trees and shrubs, at a site in the Brazilian Amazon (Taylor et al., in prep).
- The high temperature carbon compensation point of sun leaves is higher in isoprene emitting species than in non-emitting species, so it is to be expected that other aspects of heat tolerance will be similarly affected by isoprenes (https://doi.org/10.1111/pce.13564).
-  Within species, isporene production scales with light/ T

**Leaf thermal tolerance (Tcrit/ T50)**
- define.
Typical values ~45-50$^\circ$C, varying somewhat 
across latitude/climate (O’sullivan et al. 2017),
with elevation [@feeley_thermal_2020;  Slot unpublished data],
with leaf traits [Sastry et al. 2018, @zhang_photosynthetic_2012].

@@

**Differences across forest vertical gradients in biophysical conditions, plant traits, and metabolism and its thermal responses scale up to affect whole-tree ecology in several ways.** First, vertical gradients in the biophysical environment shape which plant strategies, or sets of traits, are competitive in understory versus canopy conditions. 
-	*Spring and fall leaf phenology in temperate deciduous forests (Augsburger). 	Mechanism: more moderate conditions (T, wind) in understory (REF).*
*Among species that can be deciduous, greater proportion of deciduous individuals in larger size classes (Condit et al. 2000). One potential mechanism is lower temperature/hydraulic stress in understory*
- something on isoprene production?
